\vspace{-15mm}\fontsize{21pt}{10pt}\selectfont\textbf{Wordy Manuscript Title: \\Even Wordier Manuscript Subtitle\thanks{Draft manuscript. Please do not cite without the authors' permission.}}

### 2022-09-28 sections/discussion.tex
Change: Updates to figure 7 text

  Future work building upon our current results could enhance the network inference process in multiple ways.
  The current analyses make use of one fecal microbiome transplant dataset with healthy and autistic samples, three mock community datasets, and several datasets generated by two synthetic interaction methods.
  Incorporating datasets from a broad spectrum of biomes with varying microbial distributions into \ac{micone} will likely increase the robustness and generalizability of the results from these analyses.

>>>>

  Future work building upon our current results could enhance the network inference process in multiple ways.
  The current analyses make use of one fecal microbiome transplant dataset with healthy and ASD samples, three mock community datasets, and several datasets generated by two synthetic interaction methods.
  Incorporating datasets from a broad spectrum of biomes with varying microbial distributions into \ac{micone} will likely increase the robustness and generalizability of the results from these analyses.

### 2022-09-28 sections/methods.tex
Change: Updates to figure 7 text

  \subsection*{16S rRNA sequencing datasets}
  \vspace{-5mm}
  This study utilized two kinds of 16S rRNA sequencing datasets: biological datasets and mock datasets.
  Biological datasets are collections of sequencing reads obtained from naturally occurring microbial community samples.
  The current analysis used stool samples from a fecal microbiome transplant study of autism~\cite{Kang2017} as biological datasets.
  This dataset was chosen because the sequences were easily accessible on Qiita~\cite{qiita} and optimally pre-processed according to the \ac{emp}~\cite{Thompson2017} protocol and can be used directly as input to the \ac{micone} pipeline.
  The study was composed of multiple sequencing runs.
  The runs that contained paired-end reads (run 2 (10M reads), run 3 (750K reads) and run 4 (16M reads)), were downloaded from Qiita~\cite{qiita} (study ID 10532) and used as input sequences for the \ac{micone} pipeline.
  Sequences from both control (212 samples including neurotypical and donors) and autistic (126 samples) patients were included in the analyses, but the groups were analyzed independently in the pipeline after the \ac{op} step.
  All the network analyses in the study, unless explicitly mentioned, are performed indeed on the healthy and autistic samples in the fecal microbiome transplant study.
  However, all other analyses (preceding the NI step) are performed together on all the samples in the fecal microbiome transplant study.
  The mock community 16S datasets are experimental sequencing data obtained for artificially assembled collections of species in known proportions.
  The mock datasets used for this study, obtained from mockrobiota~\cite{Bokulich2016}, are labeled mock4, mock12, and mock16.
  The mock4 community is composed of 21 bacterial strains.
  Two replicate samples from mock4 contain all species in equal abundances, and two additional replicate samples contain the same species in unequal abundances.
  The mock12 community is composed of 27 bacterial strains that include closely related taxa with some pairs having only one to two nucleotide differences from one another.
  The mock16 community is composed of 49 bacteria and 10 Archaea, all represented in equal amounts.

>>>>

  \subsection*{16S rRNA sequencing datasets}
  \vspace{-5mm}
  This study utilized two kinds of 16S rRNA sequencing datasets: biological datasets and mock datasets.
  Biological datasets are collections of sequencing reads obtained from naturally occurring microbial community samples.
  The current analysis used stool samples from a fecal microbiome transplant study of autism~\cite{Kang2017} as biological datasets.
  This dataset was chosen because the sequences were easily accessible on Qiita~\cite{qiita} and optimally pre-processed according to the \ac{emp}~\cite{Thompson2017} protocol and can be used directly as input to the \ac{micone} pipeline.
  The study was composed of multiple sequencing runs.
  The runs that contained paired-end reads (run 2 (10M reads), run 3 (750K reads) and run 4 (16M reads)), were downloaded from Qiita~\cite{qiita} (study ID 10532) and used as input sequences for the \ac{micone} pipeline.
  Sequences from both control (212 samples including neurotypical and donors) and autism spectrum disorder (ASD) (126 samples) patients were included in the analyses, but the groups were analyzed independently in the pipeline after the \ac{op} step.
  All the network analyses in the study, unless explicitly mentioned, are performed indeed on the healthy and ASD samples in the fecal microbiome transplant study.
  However, all other analyses (preceding the NI step) are performed together on all the samples in the fecal microbiome transplant study.
  The mock community 16S datasets are experimental sequencing data obtained for artificially assembled collections of species in known proportions.
  The mock datasets used for this study, obtained from mockrobiota~\cite{Bokulich2016}, are labeled mock4, mock12, and mock16.
  The mock4 community is composed of 21 bacterial strains.
  Two replicate samples from mock4 contain all species in equal abundances, and two additional replicate samples contain the same species in unequal abundances.
  The mock12 community is composed of 27 bacterial strains that include closely related taxa with some pairs having only one to two nucleotide differences from one another.
  The mock16 community is composed of 49 bacteria and 10 Archaea, all represented in equal amounts.

### 2022-09-28 sections/results.tex
Change: Updates to figure 7 text

  Our analysis involves two types of data: The first type consists of sets of 16S sequencing data from samples of human stool microbiomes from a fecal microbiome transplant study of autism~\cite{Kang2017}.
  The second type is a collection of datasets synthetically or artificially created for the specific goal of evaluating computational analysis tools.
  In particular, in order to  compare how well each step in \ac{micone} best captures the underlying data, we use both mock data (labeled mock4, mock12, and mock16) from mockrobiota~\cite{Bokulich2016} and synthetic networks generated using the NorTA~\cite{Kurtz2015} and seqtime~\cite{Rottjers2018} approaches (See Methods).

@@

  Figure~\ref{fig:figure7} shows the co-occurrence networks inferred for the healthy subjects (control) and autistic subjects in the fecal microbiome transplant study~\cite{Kang2017} (using the default tools and parameters from Table~\ref{tab:micone_tools}).
  This figure demonstrates a typical use case of comparative analysis of networks using the \ac{micone} pipeline.
  As a consequence of using the consensus network algorithm, the final co-occurrence networks are sparse and can be visually compared and examined.
  % TODO: Statistical test for differences + Refer to original paper and create conclusions
  We observed 20 unique links in the network for control samples, 12 unique links in the network for autistic subjects and 7 edges in common between both networks.
  These unique connections can serve as potential starting points for literature surveys and further experimental confirmations.

>>>>

  Our analysis involves two types of data: The first type consists of sets of 16S sequencing data from samples of human stool microbiomes from a fecal microbiome transplant (FMT) study of autism~\cite{Kang2017}.
  The second type is a collection of datasets synthetically or artificially created for the specific goal of evaluating computational analysis tools.
  In particular, in order to  compare how well each step in \ac{micone} best captures the underlying data, we use both mock data (labeled mock4, mock12, and mock16) from mockrobiota~\cite{Bokulich2016} and synthetic networks generated using the NorTA~\cite{Kurtz2015} and seqtime~\cite{Rottjers2018} approaches (See Methods).

@@

  Figure~\ref{fig:figure7} shows the co-occurrence networks inferred for the healthy subjects (control) and subjects with autism specific disorder (ASD) in the fecal microbiome transplant study~\cite{Kang2017} (using the default tools and parameters from Table~\ref{tab:micone_tools}).
  This figure demonstrates a typical use case of comparative analysis of networks using the \ac{micone} pipeline.
  As a consequence of using the consensus network algorithm, the final co-occurrence networks are sparse and can be visually compared and examined.

  % TODO: Statistical test for differences + Refer to original paper and create conclusions

  Prevotella and Clostridiales are shown to have consistent evidence of association with ASD diagnostic status [REFS - Ho et al., 2020].
  Notably, the abundance of these genera are lowered in cases with

  We observed 20 unique links in the network for control samples, 12 unique links in the network for ASD subjects and 7 edges in common between both networks.
  These unique connections can serve as potential starting points for literature surveys and further experimental confirmations.

### 2022-09-28 sections/tables_figures.tex
Change: Updates to figure 7 text

  % \FloatBarrier
  % \newpage
  % \begin{figure}[H]
  %   \centering
  %   \includegraphics[width=\textwidth]{figure2.pdf}
  % \end{figure}
  \begin{figure}[H]
    \centering
    \caption{
      \textbf{The representative sequences generated by the different DC methods differ in their identification of sequences that are low in abundance.}
      \textbf{(A)} The average weighted UniFrac distance between the representative sequences shows that the representative sequences and their compositions are fairly identical between the methods (with the exception of Deblur (DB)).
      \textbf{(B)} The relatively larger average unweighted UniFrac distance indicates that methods differ in their identification of sequences that are lower in abundance.
      The number of \ac{otu}s or \ac{esv}s generated by the respective methods are provided in the parenthesis next to the names.
      The data used for the analysis in \textbf{(A, B)} were all the samples (healthy and autistic) from the fecal microbiome transplant (FMT) dataset~\cite{Kang2017}.
      \textbf{(C, D)} The distributions of the average weighted and unweighted UniFrac distance between the expected sequence profile and the predicted sequence profile in the mock datasets.
      The distributions of the average weighted UniFrac distance show that de novo (DN) and open reference (OR) were the best-performing methods in most of the datasets, while they are worst performing methods under the unweighted UniFrac metric.
      The good performance of dada2 (D2) under both distance metrics combined with its approach of identifying \ac{esv}s using de novo methods, prompts us to propose it as the default method for the DC step.
      The data used for the analysis in \textbf{(C, D)} were the mock datasets from mockrobiota~\cite{Bokulich2016}.
    }
    \label{fig:figure2}
  \end{figure}

  % \FloatBarrier
  % \newpage
  % \begin{figure}[H]
  %   \centering
  %   \includegraphics[width=\textwidth]{figure3.pdf}
  % \end{figure}
  \begin{figure}[H]
    \centering
    \caption{
      \textbf{Taxonomic reference databases vary in terms of their taxonomy assignments beyond the Order level.}
      \textbf{(A)} The assignments of the top 50 representative sequences to their respective taxonomies using the three different reference databases.
      This result illustrates how the same sequences are assigned to different Genus under different databases.
      A significant portion of the representative sequences are assigned to an ``unknown'' Genus in two of three databases (\ac{gg} and \ac{ncbi}).
      The number of assigned genera for each database are displayed at the top of each column.
      \textbf{(B)} The number of representative sequences assigned to the same taxonomic label when using different reference databases, shown for the top 100 sequences.
      The number of mismatches are fewer at higher taxonomic levels, but, even at the Order level there exists greater than 51\% of mismatches, demonstrating the poor agreement in taxonomic labels assigned by the different databases.
      The data used for the analysis in \textbf{(A, B)} were all the samples (healthy and autistic) from the FMT dataset.
      \textbf{(C)} The Bray-Curtis dissimilarity between the expected taxonomy profile and calculated taxonomy profile in the mock datasets shows that there is no singular best choice of database for every dataset, as all the databases have similar performance.
      The \ac{gg} database and the Naive Bayes classifier are chosen as the defaults for the TA step of \ac{micone}.
      The datasets used for the analysis in \textbf{(C)} were the mock datasets from mockrobiota.
    }
    \label{fig:figure3}
  \end{figure}

@@

  % \FloatBarrier
  % \newpage
  % \begin{figure}[H]
  %   \centering
  %   \includegraphics[width=1.0\linewidth]{figure7.pdf}
  % \end{figure}
  \begin{figure}[H]
    \centering
    \caption{
      \textbf{Comparison of networks generated from autistic and control samples using the \ac{micone} pipeline}.
      The networks for the control (left) and autistic (right) samples in the FMT dataset were generated using the default tools and parameters recommended by the \ac{micone} pipeline.
      There are 20 unique links in the network for control samples, 12 unique links in the network for autistic subjects and 7 edges in common between both networks.
    }
    \label{fig:figure7}
  \end{figure}

>>>>

  % \FloatBarrier
  % \newpage
  % \begin{figure}[H]
  %   \centering
  %   \includegraphics[width=\textwidth]{figure2.pdf}
  % \end{figure}
  \begin{figure}[H]
    \centering
    \caption{
      \textbf{The representative sequences generated by the different DC methods differ in their identification of sequences that are low in abundance.}
      \textbf{(A)} The average weighted UniFrac distance between the representative sequences shows that the representative sequences and their compositions are fairly identical between the methods (with the exception of Deblur (DB)).
      \textbf{(B)} The relatively larger average unweighted UniFrac distance indicates that methods differ in their identification of sequences that are lower in abundance.
      The number of \ac{otu}s or \ac{esv}s generated by the respective methods are provided in the parenthesis next to the names.
      The data used for the analysis in \textbf{(A, B)} were all the samples from the fecal microbiome transplant (FMT) dataset~\cite{Kang2017}, containing both healthy subjects and subjects with autism spectrum disorder (ASD).
      \textbf{(C, D)} The distributions of the average weighted and unweighted UniFrac distance between the expected sequence profile and the predicted sequence profile in the mock datasets.
      The distributions of the average weighted UniFrac distance show that de novo (DN) and open reference (OR) were the best-performing methods in most of the datasets, while they are worst performing methods under the unweighted UniFrac metric.
      The good performance of dada2 (D2) under both distance metrics combined with its approach of identifying \ac{esv}s using de novo methods, prompts us to propose it as the default method for the DC step.
      The data used for the analysis in \textbf{(C, D)} were the mock datasets from mockrobiota~\cite{Bokulich2016}.
    }
    \label{fig:figure2}
  \end{figure}

  % \FloatBarrier
  % \newpage
  % \begin{figure}[H]
  %   \centering
  %   \includegraphics[width=\textwidth]{figure3.pdf}
  % \end{figure}
  \begin{figure}[H]
    \centering
    \caption{
      \textbf{Taxonomic reference databases vary in terms of their taxonomy assignments beyond the Order level.}
      \textbf{(A)} The assignments of the top 50 representative sequences to their respective taxonomies using the three different reference databases.
      This result illustrates how the same sequences are assigned to different Genus under different databases.
      A significant portion of the representative sequences are assigned to an ``unknown'' Genus in two of three databases (\ac{gg} and \ac{ncbi}).
      The number of assigned genera for each database are displayed at the top of each column.
      \textbf{(B)} The number of representative sequences assigned to the same taxonomic label when using different reference databases, shown for the top 100 sequences.
      The number of mismatches are fewer at higher taxonomic levels, but, even at the Order level there exists greater than 51\% of mismatches, demonstrating the poor agreement in taxonomic labels assigned by the different databases.
      The data used for the analysis in \textbf{(A, B)} were all the samples (healthy and ASD) from the FMT dataset.
      \textbf{(C)} The Bray-Curtis dissimilarity between the expected taxonomy profile and calculated taxonomy profile in the mock datasets shows that there is no singular best choice of database for every dataset, as all the databases have similar performance.
      The \ac{gg} database and the Naive Bayes classifier are chosen as the defaults for the TA step of \ac{micone}.
      The datasets used for the analysis in \textbf{(C)} were the mock datasets from mockrobiota.
    }
    \label{fig:figure3}
  \end{figure}

@@

  % \FloatBarrier
  % \newpage
  % \begin{figure}[H]
  %   \centering
  %   \includegraphics[width=1.0\linewidth]{figure7.pdf}
  % \end{figure}
  \begin{figure}[H]
    \centering
    \caption{
      \textbf{Comparison of networks generated from control and ASD samples using the \ac{micone} pipeline}.
      The networks for the control (left) and ASD (right) samples in the FMT dataset were generated using the default tools and parameters recommended by the \ac{micone} pipeline.
      There are 20 unique links in the network for control samples, 12 unique links in the network for ASD subjects and 7 edges in common between both networks.
    }
    \label{fig:figure7}
  \end{figure}
